---
pdf_document: default
author: "Soh Li Ying"
date: '`r Sys.Date()`'
output:
  pdf_document: default
title: "Week-5: Code-along"
---

  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# II. Code to edit and execute using the Code-along.Rmd file 
<br>

## A. Writing a function
<br>

### 1. Write a function to print a "Hello" message (Slide #14)

```{r, eval=TRUE,echo=TRUE}
say_hello_to <- function(name) {
  print(paste0("Hello ", name, "!"))
}

```
<br>

### 2. Function call with different input names (Slide #15)

```{r, eval=TRUE,echo=TRUE}
say_hello_to('Kashif')
say_hello_to('Zach')


```
<br>

### 3. typeof primitive functions (Slide #16)

```{r, eval=TRUE,echo=TRUE}
typeof(`+`)

```
<br>

### 4. typeof user-defined functions (Slide #17)

```{r, eval=TRUE,echo=TRUE}
typeof(say_hello_to)

```
<br>

### 5. Function to calculate mean of a sample (Slide #19)

```{r, eval=TRUE,echo=TRUE}
mean(rnorm(1))

```
<br>

### 6. Test your function (Slide #22)

```{r, eval=TRUE,echo=TRUE}
calc_sample_mean <- function(n) mean(rnorm(n))

```
<br>
```{r, eval=TRUE,echo=TRUE}
calc_sample_mean(2000)

```
<br>

### 7. Customizing the function to suit input (Slide #23)

```{r, eval=TRUE,echo=TRUE}
library(tidyverse)

#create a vector to test our function
sample_tibble <- tibble(sample_sizes = 
                          c(100, 300, 3000))

#using group_by to group unique entries, 
# allowing calc_sample_mean
sample_tibble %>%
  group_by(sample_sizes) %>%
  mutate(sample_means =
           calc_sample_mean(sample_sizes))


```
<br>

### 8. Setting defaults (Slide #25)

```{r, eval=TRUE,echo=TRUE}
# First define the function
calc_sample_mean <- function(sample_size, 
                             our_mean = 0, 
                             our_sd = 1) {
  
sample <- rnorm(sample_size,
                mean = our_mean,
                sd = our_sd)
mean(sample)
}

# Call the function
calc_sample_mean(sample_size = 10)

  
  
```
<br>

### 9. Different input combinations (Slide #26)

```{r, eval=TRUE,echo=TRUE}
calc_sample_mean(10, our_sd = 2)

calc_sample_mean(10, our_mean = 10)

calc_sample_mean(10, 6, 2)

```
<br>

### 10. Different input combinations (Slide #27)

```{r, eval=TRUE,echo=TRUE,error=TRUE} 
# set error=TRUE to see the error message in the output
calc_sample_mean(our_mean = 5)
```
<br>

### 11. Some more examples (Slide #28)

```{r, eval=TRUE,echo=TRUE,error=FALSE} 
# Add 2 to input x
add_two <- function(x) {
  x+2
}

add_two

add_two(-34)
```
<br>

## B. Scoping
<br>

### 12. Multiple assignment of z (Slide #36)

```{r, eval=TRUE,echo=TRUE}
# Initialize z
z <- 1
sprintf("The value assigned to z outside the function is %d", z)

#declare a function, passing value of 2 for z
addition_of_three <- function (z = 2) {
  #reassigning z
  z <-3
  return(z+3)
}

addition_of_three()

```
<br>

### 13. Multiple assignment of z (Slide #37)

```{r, eval=TRUE,echo=TRUE}
# Initialize z
z <- 1

#declare a function, passing value of 2 for z
addition_of_three <- function(z = 2){
  #reassigning z
  z <- 3
  return(z+3)
}

#another reassingment of z 
addition_of_three(z = 4)

#Accessing z outside of the function
sprintf("The final value of z after reassigning it to a different value inside the function is %d", z)
```
<br>