---
pdf_document: default
author: "Soh Li Ying"
date: "`r Sys.Date()`"
output:
  pdf_document: default
title: "Week-4: Code-along"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# II. Code to edit and execute using the Code-along.Rmd file 
<br>

## A. Data Wrangling
<br>

### 1. Loading packages (Slide #16)

```{r, eval=TRUE,echo=TRUE}
# Load package tidyverse
library(tidyverse)

```

### 2. Loading data-set (Slide #16)

```{r, eval=TRUE,echo=TRUE}
# Read data from the hotels.csv file and assign it to a variable named, "hotels"
hotels <- read_csv("hotels.csv")

```

### 3. List names of the variables in the data-set (Slide #19)

```{r, eval=TRUE,echo=TRUE}
# List names of variables
names(hotels)

```


### 4. Glimpse of contents of the data-set (Slide #20)

```{r, eval=TRUE,echo=TRUE}
# Glimpse of contents
glimpse(hotels)

```
<br>

## B. Choosing rows or columns
<br>

### 5. Select a single column (Slide #24)

```{r, eval=TRUE,echo=TRUE}
# Selecting only "lead_time" column
select(hotels, lead_time)

```

### 6. Select multiple columns (Slide #25)

```{r, eval=TRUE,echo=TRUE}
# Selecting columns lead_time, country, and reserved_room_type
select(hotels, lead_time,country,reserved_room_type)

```

### 7. Arrange entries of a column (Slide #28)

```{r, eval=TRUE,echo=TRUE}
# Arrange entries of lead_time column
arrange(hotels, lead_time)

```

### 8. Arrange entries of a column in the descending order (Slide #30)

```{r, eval=TRUE,echo=TRUE}
# Arrange entries of lead_time in descending order
arrange(hotels, desc(lead_time))

```

### 9. Select columns and arrange the entries of a column (Slide #31)

```{r, eval=TRUE,echo=TRUE}
# Select lead_time & country column, and arrange entries
arrange(select(hotels, lead_time,country), desc(lead_time))

```

### 10. Select columns and arrange the entries of a column using the pipe operator (Slide #37)

```{r, eval=TRUE,echo=TRUE}
# Select lead_time & country columns and arrange the entries in decreasing orded using the pipe operator
hotels %>% select(lead_time, country) %>% arrange(desc(lead_time)) 

```

### 11. Pick rows matching a condition (Slide #44)

```{r, eval=TRUE,echo=TRUE}
# Pick rows where # of children >= 1 
hotels %>% filter(children >= 1) %>% select(hotel, children)

```

### 12. Pick rows matching multiple conditions (Slide #46)

```{r, eval=TRUE,echo=TRUE}
# Pick rows matching conditions of both children >=1 and only "City Hotel"
hotels %>% filter(children >= 1, hotel == "City Hotel") %>% select(hotel, children)

```

### 13. Non-conditional selection of rows: sequence of indices (Slide #49)

```{r, eval=TRUE,echo=TRUE}
# Using a sequence of indices
hotels %>% slice(1:5)

```

### 14. Non-conditional selection of rows: non-consecutive/specific indices (Slide #50)

```{r, eval=TRUE,echo=TRUE}
# non-consecutive/specific indices
hotels %>% slice(1,3,5)

```

### 15. Pick unique rows using distinct() (Slide #52)

```{r, eval=TRUE,echo=TRUE}
# Picking unique rows
hotels %>% distinct(hotel)

```
<br>

## C. Creating new columns
<br>

### 16. Creating a single column with mutate() (Slide #56)

```{r, eval=TRUE,echo=TRUE}
# Creating a single column
hotels %>% mutate(little_ones = children + babies) %>% select(hotel, little_ones,children,babies)

```

### 17. Creating multiple columns with mutate() (Slide #58)

```{r, eval=TRUE,echo=TRUE}
# Creating multiple columns 
hotels %>%
  mutate(little_ones = children + babies, 
         average_little_ones = mean(little_ones)) %>% select(hotel, little_ones,average_little_ones)

```
<br>

## D. More operations with examples
<br>

### 18. count() to get frequencies (Slide #60)

```{r, eval=TRUE,echo=TRUE}
# count()
hotels %>% count(market_segment)

```

### 19. count() to get frequencies with sorting of count (Slide #61)

```{r, eval=TRUE,echo=TRUE}
# Frequencies w sorting of count 
hotels %>%
  count(market_segment, sort = TRUE)

```

### 20. count() multiple variables (Slide #62)

```{r, eval=TRUE,echo=TRUE}
# count() multiple variables 
hotels %>% count(hotel, market_segment)

```

### 21. summarise() for summary statistics (Slide #63)

```{r, eval=TRUE,echo=TRUE}
# specific summary stats 
hotels %>% summarise(mean_adr = mean(adr))

```

### 22. summarise() by using group_by to find mean (Slide #64)

```{r, eval=TRUE,echo=TRUE}
# summarise() by using group_by to find mean
hotels %>% 
  group_by(hotel) %>%
  summarise(mean_adr = mean(adr))

```

### 23. summarise() by using group_by to get count (Slide #65)

```{r, eval=TRUE,echo=TRUE}
# summarise() by using group_by to get count
hotels %>%
  group_by(hotel) %>%
  summarise(count = n())

```

### 24. summarise() for multiple summary statistics (Slide #67)

```{r, eval=TRUE,echo=TRUE}
# summarise() for multiple summary statistics
hotels %>% 
  summarise(min_adr = min(adr),
            mean_adr = mean(adr),
            median_adr = median(adr),
            max_adr = max(adr))

```

### 25. select(), slice() and arrange() (Slide #68)

```{r, eval=TRUE,echo=TRUE}
# select(), slice() and arrange()
hotels %>% 
  select(hotel, lead_time) %>%
  slice(1:5) %>% 
  arrange(lead_time)

```

### 26. select(), arrange() and slice() (Slide #69)

```{r, eval=TRUE,echo=TRUE}
# select(), arrange() and slice()
hotels %>%
  select(hotel, lead_time) %>%
  arrange(lead_time) %>%
  slice(1:5)

```

### 27. filter() to select rows based on conditions (Slide #73)

```{r, eval=TRUE,echo=TRUE}
# filter() to select rows that has "City Hotel" as hotels
hotels %>% 
  filter (hotel == "City Hotel")

```

### 28. filter() to select rows based on complicated conditions (Slide #74)

```{r, eval=TRUE,echo=TRUE}
# filter() to select rows where adults = 1, children >= 1 or babies >= 1 
hotels %>% 
  filter(adults == 1,
         children >=1 | babies >= 1) %>%
  select(adults, babies, children)

```

### 29. count() and arrange() (Slide #76)

```{r, eval=TRUE,echo=TRUE}
# count() and arrange()
hotels %>% 
  count(market_segment) %>%
  arrange(desc(n))

```

### 30. mutate(), select() and arrange() (Slide #77)

```{r, eval=TRUE,echo=TRUE}
# mutate(), select() and arrange()
hotels %>%
  mutate(little_ones = children + babies) %>%
  select(children, babies, little_ones) %>%
  arrange(desc(little_ones))

```


### 31. mutate(), filter() and select() (Slide #78)

```{r, eval=TRUE,echo=TRUE}
# mutate(), filter() and select()
hotels %>%
  mutate(little_ones = children + babies) %>%
  filter(little_ones >= 1, hotel == "City Hotel") %>%
  select(hotel, little_ones)

```