---
title: "Challenge-6"
author: "Soh Li Ying"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  'pdf_document: default': default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Questions
<br>

#### Question-1: Countdown Blastoff (While Loop)

Create a program that simulates a rocket launch countdown using a while loop. Start from 10 and countdown to "Blastoff!" with a one-second delay between each countdown number. Print a message when the rocket launches.

**Hint:** You may want to use `cat` command to print the countdown and `Sys.sleep` for incorporating the delay

**Solutions:**

```{r, eval=T,echo=TRUE}
countdown <- 10

# While loop to countdown
while (countdown >= 0) {
  if (countdown == 0) {
    cat("\nBlastoff!")
  } else {
    cat(countdown, "...")
  }
  # Sleep for one second
  Sys.sleep(1)
  # Iterate from the back 
  countdown <- countdown - 1
}

```
<br>

#### Question-2: Word Reverser (for Loop)

Develop a program that takes a user-entered word and uses a while loop to print the word's characters in reverse order. For example, if the user enters "hello," the program should print "olleh."

**Hint:** You may want to use `substr` command to access each character of the input word, and `paste` command to join the reversed letters one at a time

**Solutions:**

```{r, eval=T,echo=TRUE}
# Prompt user to input a word
user_input <- "butterly"

# Track position of character
pos <- nchar(user_input)

# Initialise an empty string to store the reversed word
reversed_word <- character()

# While loop
while (pos >= 1) {
  # Accessing each character of the input word in current position
  char <- substr(user_input, pos, pos)
  # Append characters to reversed word
  reversed_word <- paste(reversed_word, char, sep = "")
  # Work backwards to the last letter
  pos <- pos - 1
}

# Print the reversed word
cat("Reversed word: ", reversed_word, "\n")

```
<br>

