---
title: "R Notes"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<style>
p.comment {
background-color: #DBDBDB;
padding: 10px;
border: 1px solid black;
margin-left: 25px;
border-radius: 5px;
font-style: italic;
}

</style>

##Installing, running and importing data
**library(*package_name*)** - load R package 


## Creating, listing and deleting objects in memory:
**<- / ->** Creating an object with the "assign" operator

**ls()**: lists the object in memory 

**pat = **: (pattern) lists only the objects which CONTAINS a given character
```{r, eval=FALSE,echo=TRUE}
ls(pat = "m")
``` 
To restrict list to names that STARTS with the character
```{r, eval=FALSE,echo=TRUE}
ls(pat = "^m")
```

## List of useful commands:
**read_csv**("*name_of_file.csv*")** - Import dataset into R.

**? / help() / help("")** -  Find out more about command.  
(The last one MUST be used to access non-conventional characters such as *)

**glimpse(*data_file_name*)** - Display all the columns of the variable stacked one below another.

**head(*data_file_name*)** -  Display the first few rows of the data-set to get an overview of its structure.

**addNA** – Create a new factor/vector with an additional category representing NA (missing values).  

**$** – Operator to extract a specific portion of data.

**%>%** – Pipe operator that passes the output of one function as an input to another.  
**filter(*row_attributes*)** – Subset rows using column values.  
**select(*col_names*)** – Subset column using their names/types.  
```{r, eval=FALSE,echo=TRUE}
dataset_name %>% filter(coloumn_name%in%dataset) %>% select(coloumns_required)
```

## Creating vectors & lists: 
**list()** – Create a list object containing function arguments.  
**as.list() – Convert an object into a list

## Mathematical functions:
**abs()** – Compute the absolute value of a numeric data object.
**aggregate** – Compute summary statistics of subgroups of a data set.
**ave()** – Calculate averages.  
**ceiling()** – Round numeric up to the next higher integer. 
**IQR()** – Compute the Interquartile Range.  
**sum()** – Compute the sum of a numeric input vector.  


## Converting datatypes: 
**as.data.frame** – Coerce object to data frame.  
**as.double/as.numeric/as.str** – Convert a data object to double class/ numeric class/ str class.  
**as.factor** – Convert a data object to the class factor.  
**as.function** – Convert an object to a function.  
**as.name** – Convert character string to name class object.    

## Strings:
**str_c [stringr]** – Combine several inputs into a single character vector.  
**str_count [stringr]** – Count number of matches in a character string.  
**str_detect [stringr]** – Detect presence or absence of pattern in character string.  
**str_dup [stringr]** – Duplicate and concatenate strings within a character vector.  
**str_extract [stringr]** – Extract matching patterns from character string.  
**str_length [stringr]** – Return the length of a string.  
**str_locate [stringr]** – Locate the first position of patterns in a character string.  
**str_locate_all [stringr]** – Locate the position of patterns in a character string.  
**str_order [stringr]** – Order a vector of character strings.  
**str_remove [stringr]** – Remove FIRST matched patterns in a character string.  
**str_remove_all [stringr]** – Remove ALL matched patterns in a character string.  
**str_replace [stringr]** – Replace FIRST matched patterns in character strings.  
**str_replace_all [stringr]** – Replace ALL matched patterns in character strings.  
**str_replace_na [stringr]** – Replace NA values in a character string.  
**str_sort [stringr]**v – Sort a vector of character strings.  
**vstr_split [stringr]** – Split up a string into pieces.  
**str_split_fixed** [stringr] – Split up a string into a fixed amount of pieces.  
**str_trim [stringr]** – Trim whitespace from a character string.  
**str_which [stringr]** – Identify positions in a character string matching a pattern.  
**str_wrap[stringr]** – Wrap a character string into nicely formatted paragraphs.  
**str** – Compactly display the structure of a data object. 


##Visualising data:

### Graph plots:
**abline** – Add straight lines to plot.
**box** – Draw a border around a plot.  


### Other plots: 
**barplot** – Create a barplot.  
**boxplot** – Create a boxplot.  

## Beautify tables/arrays
**addmargins** – Put margins on tables or arrays.  


## Other functions: 
**all** – Check whether all values of a logical vector are TRUE.  
**any** – Check whether any values of a logical vector are TRUE.  
**apply** – Apply function by rows or columns of a data frame.  
**approx** – Return a list of values that linearly interpolate given data points.  
**apropos** – Return character vector with names of objects that contain the input.  
**arrange** – Order data frames and tibbles.  
**casefold** – Translate character to lower or upper case.  
**chartr(*old*, *new*, *x*)** – Character translation/replacement  
cat("- *old* = A character vector specifying characters to be replaces")  
cat("-*new* = A character vector specifying the corresponding replacement characters")  
cat("- *x* = A character vector in which replacement will occur")  
**floor** – Round numeric down to the next lower integer.  

## Creating loops & iterations: 
**break** – Break for-loop in R.  
**if_else** [dplyr] – Conditionally select an element.  
**ifelse** – Select elements conditionally.  




clip – Specify the clipping region of a graphic.
coalesce [dplyr] – Replace missing values.
combine [dplyr] – Combine values.
digits [TeachingDemos] – Add straight lines to plot.
dnorm – Create standard normal distribution.
download.file – Download file from internet.
duplicated – Determine duplicate elements.
eval – Evaluate an expression and return the result.
everything [tidyselect] – Select all variables of a data set.
exists – Check whether an object is defined in the R environment.
exp – Compute exponentials.
find – Return location where objects of a given name can be found.

geom_bar – Create a barplot.
geom_boxplot [ggplot2] – Create a boxplot.
geom_curve [ggplot2] – Add curve to ggplot2 plot.
geom_histogram [ggplot2] – Draw a histogram.
geom_line [ggplot2] – Draw a line graph.
geom_segment [ggplot2] – Add line segment to ggplot2 plot.
get – Search and call a data object.
hist – Create histogram.
identity – Return a data object.


is.data.frame – Test if object is a data frame.
is.double – Test whether a data object has the double class.
is.function – Test if an object is a function.
is.na – Return a logical vector or matrix indicating which elements are missing.
is.name – Test whether a data object has the class name.
is.nan – Return a logical vector or matrix indicating which elements are not a number.
is.null – Return a logical value indicating whether a data object is of the data class NULL.
isprime [matlab] – Check for prime number.
lapply – Apply function to all list elements.
length – Return the length of data objects such as vectors or lists.
load – Load RData workspace file into R.
log – Compute logarithms of numeric values.
log10 – Compute logarithms of numeric values with base of 10.
log2 – Compute logarithms of numeric values with base of 2.
ls – Return the names of data objects in an environment.
match – Return position of first match between two data objects.
max – Compute the maximum value of a vector or column.
mean – Compute the arithmetic mean.
median – Compute the median.
merge – Merge data frames.K
na.omit – Remove incomplete cases of a data object.
na_if [dplyr] – Replace NA values.
names – Return or set the names of a data object.
nchar – Return the number of elements (e.g. letters) of a character object.
ncol – Return the number of columns of a matrix or data frame.
near [dplyr] – Compare numeric vectors.
next – Skip certain parts of for-loop in R.
norm – Compute the norm of a matrix.
nrow – Return the number of rows of a matrix or data frame.
nth [dplyr] – Extract certain values from vector.
ntile [dplyr] – Rank a vector.
order – Return position of each input element in ascending or descending order.
paste – Concatenate inputs to string.
paste0 – Concatenate inputs to string with given separator.
percent [formattable] – Format number as percentage.
percent [scales] – Format number as percentage.
percent_rank [dplyr] – Rank a vector.
plnorm – Return corresponding value of log normal CDF.
plogis – Return corresponding value of logistic CDF.
plot – Draw a scatterplot or density plot.
pnorm – Return value of distribution function.
polygon – Draw a polygon to a plot.
ppois – Return value of poisson cumulative distribution function.
print – Return data object to the R (or RStudio) console.
pull [dplyr] – Extract columns of data frame or tibble.
punif – Return corresponding value of uniform CDF.
qqline – Add theoretical line to QQplot.
qqnorm – Create a normal QQplot.
qqplot – Create QQplot of two data sets.
quit – Terminate an R session.
range – Get the minimum and maximum of a data object.
rank – Return ranking position of input.
read.xlsx – Read Excel File in R.
read_excel [readxl Package] – Read Excel File in R.
remove – Clear specific data object from R workspace.
rename [dplyr] – Rename variables of data frame or tibble.
reorder – Reorder Levels of Factor.
rep – Replicate elements of vectors and lists.
replace – Exchanges values in data object.
replicate – Perform a repeated evaluation of an expression.
return – Return output of user-defined R function.
rm – Clear specific data object from R workspace.
round – Round numeric to specified number of decimal places.
row.names – Get or set row names of data frame or matrix.
row – Return row indices and labels.
rowMeans – Compute the mean of each row of a numeric data frame, matrix or array.
rowMedians [robustbase] – Compute the median of each row of a numeric matrix.
rowSums – Compute the sum of each row of a numeric data frame, matrix or array.
sapply – Apply function to all list elements and convert output to vector.
save.image – Save global R environment as RData workspace file.
save – Save R data objects as RData workspace file.
scale – Standardize data.
sd – Compute standard deviation.
search – Return a list of attached packages and objects.
select [dplyr] – Select variables of data frame or tibble.
seq – Generate sequence of numeric values.
signif – Round numeric to specified number of digits.
slice [dplyr] – Extract certain rows of data set.
solve – Solve system of equations.
sort.int – Return input in ascending or descending order.
sort.list – Return position of each element of a single input.
sort – Return input in ascending or descending order.
split – Divide data into groups.
sqrt – Compute the square root of a numeric data object.
stat_qq [ggplot2] – Create a QQplot.
stat_qq_line [ggplot2] – Add theoretical line to QQplot.
std.error [plotrix] – Compute standard error in R.
stop – Generate error message and stop executing the current R code.
stopifnot – Check if all expressions are TRUE.

structure – Set additional attributes of a data object.
sub – Replace first match in character string.
substr – Extract or replace a substring in a character vector.
substring – Extract or replace a substring in a character vector.

summary – Compute summary statistics of data and model objects.
tail – Return the last part of a data object.
tapply – Apply function to multiple data groups.
tolower – Translate character to lower case.
toupper – Translate character to upper case.
trunc – Cut off decimal places of a numeric input.
type.convert – Convert data object to appropriate class.
unique – Remove duplicates from data objects.
var – Compute sample variance.
which – Return indices of certain values based on logical condition.
which.max – Determine location of first maximum.
which.min – Determine location of first minimum.
with – Evaluates expression in environment constructed based on data frame.
within – Evaluates expression in environment and modify data frame.
write_clip – Copy data to the clipboard.

## Other useful commands 
```{r, eval=FALSE,echo=TRUE}
<br>: Add on the next line to create an empty line 
``` 
**(2 double spaces)**: Next line


## Keyboard shortcuts:
**Ctrl + Shift + K**: Output doc in format specified in your YAML header 

**Ctrl + Shift + F10**: Restart R session to start w a clean env

**Ctrl + Alt + P**: Run all chunks above your current chunk

**Ctrl + Enter**: Run current line of code

**Ctrl + Alt + R**: Run all chunks

## Code Chunk Options:
**echo = FALSE**: Hide the code, but run code and produce all outputs, plots, warnings and messages

**eval = FALSE**: Show code, but do not evaluate it

**fig.show = "hide"**: Hides plots

**include = FALSE**: Runs code, but suppresses all output

**message = FALSE**: Prevent packages from printing messages when they load. This also suppress messages generated by functions.

**results = "hide"**: Hides printed output.

**warning = FALSE**: Prevents packages and functions from displaying warnings.
